\documentclass[oneside, a4paper]{memoir}
\setlength\parindent{0pt}
\setsecnumdepth{section}
\counterwithout{section}{chapter}
\title{\LaTeX{} Installation Guide}
\author{Tim Birger Tejsner}
\usepackage{csquotes}
\usepackage[english]{babel}
\usepackage{url}
\usepackage{geometry}
\usepackage{listings}
\usepackage{xcolor}

\lstset
{
    language=[LaTeX]TeX,
    breaklines=true,
    basicstyle=\ttfamily,
    keywordstyle=\color{blue},
    identifierstyle=\color{black},
    morekeywords={chapter, subsection, subsubsection, chapterstyle}
}

\begin{document}

\maketitle

This document contains a short guide to install a working \LaTeX{} environment on your system. 

\section{\LaTeX{} distribution}
First we need a \LaTeX{} distribution. This is used to compile \texttt{.tex} files into viewable \texttt{.pdf} files.

\subsection{Windows}
\begin{itemize}
\item Go to \url{http://miktex.org/download}.
\item Download the \emph{Basic MikTeX Installer}.
\item Run the installer and follow the on-screen instructions.
\item Under \enquote{Install missing packages on-the-fly} choose \enquote{yes}.
\end{itemize}

\subsection{OSX}
\begin{itemize}
\item Go to \url{https://tug.org/mactex/mactex-download.html}
\item Download \texttt{MacTeX.pkg}
\item Double-click the file to install
\end{itemize}

\subsection{Linux}
For Linux installations I recommend texlive located at \url{https://www.tug.org/texlive/}. Most Linux distributions have texlive in official repositories. For Ubuntu simply type 
\begin{quote}
\texttt{sudo apt-get install texlive-full}
\end{quote}
in a terminal window.


\section{Texmaker}
Texmaker is the editor we will use to write \LaTeX{} documents.

\begin{itemize}
\item Go to \url{http://www.xm1math.net/texmaker/download.html}
\item Install the Version appropriate for your system.
\end{itemize}

\section{Testing the installation}
Open Texmaker and create a new document. Paste the following into the editor:

\begin{quote}
\begin{lstlisting}
\documentclass[a4paper]{memoir}
\usepackage[utf8]{inputenc}
\begin{document}
Hello World!
\[ e^{i\pi} = -1 \]
\end{document}
\end{lstlisting}
\end{quote}

Save the document. Go to \enquote{Tools} and chose \enquote{Quick Build}. This should generate your \texttt{.pdf}.

\end{document}